%% start of file `template.tex'.
%% Copyright 2006-2015 Xavier Danaux (xdanaux@gmail.com).
%
% Adapted to be an Rmarkdown template by Mitchell O'Hara-Wild
% 8 February 2019
%
% This work may be distributed and/or modified under the
% conditions of the LaTeX Project Public License version 1.3c,
% available at http://www.latex-project.org/lppl/.


\documentclass[11pt,a4paper,]{moderncv}

% moderncv themes
\moderncvstyle{casual}                             % style options are 'casual' (default), 'classic', 'banking', 'oldstyle' and 'fancy'

\definecolor{color0}{rgb}{0,0,0}% black
\definecolor{color1}{HTML}{3873B3}% custom
\definecolor{color2}{rgb}{0.45,0.45,0.45}% dark grey

\usepackage[scaled=0.86]{DejaVuSansMono}

\providecommand{\tightlist}{%
	\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\def\donothing#1{#1}
\def\emaillink#1{#1}

%\nopagenumbers{}                                  % uncomment to suppress automatic page numbering for CVs longer than one page

% character encoding
%\usepackage[utf8]{inputenc}                       % if you are not using xelatex ou lualatex, replace by the encoding you are using
%\usepackage{CJKutf8}                              % if you need to use CJK to typeset your resume in Chinese, Japanese or Korean

% adjust the page margins
\usepackage[scale=0.75,footskip=60pt]{geometry}
%\setlength{\hintscolumnwidth}{3cm}                % if you want to change the width of the column with the dates
%\setlength{\makecvheadnamewidth}{10cm}            % for the 'classic' style, if you want to force the width allocated to your name and avoid line breaks. be careful though, the length is normally calculated to avoid any overlap with your personal info; use this at your own typographical risks...



% personal data
\name{}{Dr.~Deniz Akdemir}
\title{Senior Clinical Data Scientist - Statistical Genomics Consultant}


\phone[mobile]{+1 607 262 6875 (USA) - +34 674 771 859
(Spain)} % Phone number
\email{\donothing{\href{mailto:deniz.akdemir.work@gmail.com}{\nolinkurl{deniz.akdemir.work@gmail.com}}}}
 % Personal website


\social[github]{denizakdemir}



% \extrainfo{additional information}                 % optional, remove / comment the line if not wanted




% Pandoc CSL macros
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

%----------------------------------------------------------------------------------
%            content
%----------------------------------------------------------------------------------
\begin{document}
%\begin{CJK*}{UTF8}{gbsn}                          % to typeset your resume in Chinese using CJK
%-----       resume       ---------------------------------------------------------
\makecvtitle



\section{Core Skills}\label{core-skills}

\begin{itemize}
\tightlist
\item
  \textbf{Statistical Genomics:} Genomic selection, multi-trait
  methodologies, genomic prediction, GWAS, models for epistasis and
  genotype by environment interactions, optimization of plant and animal
  breeding programs.
\item
  \textbf{Statistics:} Multivariate analysis, high-dimensional data
  modeling, Bayesian methods, machine learning, deep learning, data
  science, biostatistics.
\item
  \textbf{Statistical Computation:} Proficient in R, SAS, Python, and
  C++ for statistical programming and analysis.
\item
  \textbf{Software Development:} Development of research and commercial
  grade software.
\item
  \textbf{Management and administration:} Strategic planning, project
  management, team leadership, grant/patent writing, budgeting,
  accounting.
\end{itemize}

\section{Career Summary}\label{career-summary}

\begin{itemize}
\tightlist
\item
  \textbf{Senior Clinical Data Scientist, National Marrow Donor Program,
  Minneapolis, USA (2023 - Current):} Engaged in statistical and machine
  learning analysis of stem cell transplant data, focusing on research
  into donor optimization. Applied for grants and submitted manuscripts
  to peer-reviewed journals. Wrote patent applications.
\item
  \textbf{Clinical Data Scientist, National Marrow Donor Program,
  Minneapolis, USA (2021 - 2023):} Engaged in statistical and machine
  learning analysis of stem cell transplant data, focusing on research
  into donor optimization.
\item
  \textbf{Postdoctoral Research Associate, School of Agriculture and
  Food Science, University College Dublin, Dublin, Ireland (2019 -
  2021):} Conducted research on methods for combining heterogeneous
  genomic and phenotypic datasets and prepared statistical software for
  data analysis.
\item
  \textbf{Statistical Consultant, Cornell Statistical Consulting Unit,
  Cornell University, Ithaca, NY, USA (2017 - 2019):} Provided
  statistical consulting services for researchers at Cornell University,
  including the preparation and presentation of statistics workshops.
\item
  \textbf{Postdoctoral Research Associate, Department of Plant Breeding
  and Genetics, Cornell University, Ithaca, NY, USA (2011 - 2017):}
  Focused on research developing new methodologies in genomic selection
  and prediction, mixed models, and machine learning, advising graduate
  students and preparing statistical software.
\item
  \textbf{Visiting Assistant Professor, Department of Statistics and
  Actuarial Science, University of Central Florida, Orlando, FL, USA
  (2010 - 2011):} Responsibilities included teaching Data Mining
  Methodology, Theoretical Statistics, Applied Time Series Analysis, and
  Nonparametric Statistics.
\item
  \textbf{Visiting Assistant Professor, Department of Mathematics and
  Statistics, Ohio Northern University, Ada, OH, USA (2009 - 2010):}
  Taught Statistics for Professionals, Statistics for Engineers, and
  Statistical Computing, catering to various undergraduate levels.
\end{itemize}

\section{Workshops and Training
Sessions}\label{workshops-and-training-sessions}

\begin{itemize}
\tightlist
\item
  \textbf{Genomic Assisted Breeding Workshops:} Conducted workshops in
  the USA, Ireland, Spain, and Belgium, educating the agricultural
  community on genomic technologies.
\item
  \textbf{Public Speaking:} Presented at various international
  conferences including EFI conference in Geneva and ASHI Annual Meeting
  in San Antonio.
\end{itemize}

\section{Publications}\label{publications}

\phantomsection\label{refs-fc9fcad9798ae22e92c54d59a6d1bc3c}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-akdemir2020combining}
Akdemir, D., Knox, R., \& Isidro y Sanchez, J. (2020). Combining
partially overlapping multi-omics data in databases using relationship
matrices. \emph{Frontiers in Plant Science}, \emph{11}, 947.
\url{https://doi.org/10.3389/fpls.2020.00947}

\bibitem[\citeproctext]{ref-pinho2020using}
Pinho Morais, P. P., Akdemir, D., Braatz de Andrade, L. R., Jannink,
J.-L., Fritsche-Neto, R., Borem, A., \ldots{} Granato, I. S. C. (2020).
Using public databases for genomic prediction of tropical maize lines.
\emph{Plant Breeding}, \emph{139}(4), 697--707.
\url{https://doi.org/10.1111/pbr.12827}

\bibitem[\citeproctext]{ref-akdemir2019multi}
Akdemir, D., Beavis, W., Fritsche-Neto, R., Singh, A. K., \&
Isidro-Sanchez, J. (2019). Multi-objective optimized genomic breeding
strategies for sustainable food improvement. \emph{Heredity},
\emph{122}(5), 672--683.
\url{https://doi.org/10.1038/s41598-018-38081-6}

\bibitem[\citeproctext]{ref-akdemir2019design}
Akdemir, D., \& Isidro-Sanchez, J. (2019). Design of training
populations for selective phenotyping in genomic prediction.
\emph{Scientific Reports}, \emph{9}(1), 1--15.
\url{https://doi.org/10.1038/s41598-018-38081-6}

\bibitem[\citeproctext]{ref-ikeogu2019genetic}
Ikeogu, U. N., Akdemir, D., Wolfe, M. D., Okeke, U. G., Chinedozi, A.,
Jannink, J.-L., \& Egesi, C. N. (2019). Genetic correlation, genome-wide
association and genomic prediction of portable NIRS predicted
carotenoids in cassava roots. \emph{Frontiers in Plant Science},
\emph{10}, 1570. \url{https://doi.org/10.3389/fpls.2019.01570}

\bibitem[\citeproctext]{ref-kaya2019genome}
Kaya, H. B., Akdemir, D., Lozano, R., Cetin, O., Kaya, H. S., Sahin, M.,
\ldots{} Jannink, J.-L. (2019). Genome wide association study of 5
agronomic traits in olive (olea europaea l.). \emph{Scientific Reports},
\emph{9}(1), 1--14. \url{https://doi.org/10.1038/s41598-019-55338-w}

\bibitem[\citeproctext]{ref-fristche2018accuracy}
Fristche-Neto, R., Akdemir, D., \& Jannink, J.-L. (2018). Accuracy of
genomic selection to predict maize single-crosses obtained through
different mating designs. \emph{Theoretical and Applied Genetics},
\emph{131}(5), 1153--1162.
\url{https://doi.org/10.1007/s00122-018-3068-8}

\bibitem[\citeproctext]{ref-horn2018importance}
Horn, L., Shimelis, H., Kariaga, M., Onamu, R., Wakhungu, J., Were, H.,
et al.others. (2018). Importance of cowpea production, breeding and
production constrains under dry areas in africa. \emph{RUFORUM Working
Document Series}, \emph{17}(1), 499--514.
\url{https://doi.org/10.1016/j.aoas.2020.03.002}

\bibitem[\citeproctext]{ref-okeke2018regional}
Okeke, U. G., Akdemir, D., Rabbi, I., Kulakow, P., \& Jannink, J.-L.
(2018). Regional heritability mapping provides insights into dry matter
content in african white and yellow cassava populations. \emph{The Plant
Genome}, \emph{11}(1), 170050.
\url{https://doi.org/10.3835/plantgenome2017.06.0050}

\bibitem[\citeproctext]{ref-akdemir2017locally}
Akdemir, D., Jannink, J.-L., \& Isidro-Sanchez, J. (2017). Locally
epistatic models for genome-wide prediction and association by
importance sampling. \emph{Genetics Selection Evolution}, \emph{49}(1),
1--14. \url{https://doi.org/10.1186/s12711-017-0348-8}

\bibitem[\citeproctext]{ref-isidro2017genome}
Isidro-Sanchez, J., Akdemir, D., \& Montilla-Bascon, G. (2017).
Genome-wide association analysis using r. In \emph{Oat} (pp. 189--207).
\url{https://doi.org/10.1007/978-1-4939-6682-0_14}

\bibitem[\citeproctext]{ref-okeke2017accuracies}
Okeke, U. G., Akdemir, D., Rabbi, I., Kulakow, P., \& Jannink, J.-L.
(2017). Accuracies of univariate and multivariate genomic prediction
models in african cassava. \emph{Genetics Selection Evolution},
\emph{49}(1), 1--10. \url{https://doi.org/10.1101/116301}

\bibitem[\citeproctext]{ref-ratervar2017}
Stansell, Z., Akdemir, D., \& Bjorkman. (2017). Use of a quality trait
index to increase the reliability of phenotypic evaluations.
\emph{HortScience}, \emph{52}, 1490.

\bibitem[\citeproctext]{ref-akdemir2016array}
Akdemir, D. (2016). Array normal model and incomplete array variate
observations. In \emph{Applied matrix and tensor variate data analysis}
(pp. 93--122). \url{https://doi.org/10.1007/978-4-431-55387-8_5}

\bibitem[\citeproctext]{ref-akdemir2016efficient}
Akdemir, D., \& Sanchez, J. I. (2016). Efficient breeding by genomic
mating. \emph{Frontiers in Genetics}, \emph{7}, 210.
\url{https://doi.org/10.3389/fgene.2016.00210}

\bibitem[\citeproctext]{ref-ashraf2016genomic}
Ashraf, B., Edriss, V., Akdemir, D., Autrique, E., Bonnett, D., Crossa,
J., \ldots{} Jannink, J.-L. (2016). Genomic prediction using phenotypes
from pedigreed lines with no marker data. \emph{Crop Science},
\emph{56}(3), 957--964.
\url{https://doi.org/10.2135/cropsci2015.02.0111}

\bibitem[\citeproctext]{ref-crossa2016genomic}
Crossa, J., Jarquin, D., Franco, J., Perez-Rodriguez, P., Burgueno, J.,
Saint-Pierre, C., et al.others. (2016). Genomic prediction of gene bank
wheat landraces. \emph{G3: Genes, Genomes, Genetics}, \emph{6}(7),
1819--1834.

\bibitem[\citeproctext]{ref-jafarzadeh2016breeding}
Jafarzadeh, J., Bonnett, D., Jannink, J.-L., Akdemir, D., Dreisigacker,
S., \& Sorrells, M. E. (2016). Breeding value of primary synthetic wheat
genotypes for grain yield. \emph{PloS One}, \emph{11}(9), e0162860.
\url{https://doi.org/10.1371/journal.pone.0162860}

\bibitem[\citeproctext]{ref-jafarzadehdetermining}
Jafarzadeh, J., Bonnett, D., Jannink, J.-L., Akdemir, D., \& Sorrells,
M. E. (2016). Determining the breeding value of primary synthetic wheats
for increased grain yield. \emph{PLOS ONE}, \emph{11}(9), e0162860.
\url{https://doi.org/10.1371/journal.pone.0162860}

\bibitem[\citeproctext]{ref-spindel2016genome}
Spindel, J., Begum, H., Akdemir, D., Collard, B., Redona, E., Jannink,
J., \& McCouch, S. (2016). Genome-wide prediction models that
incorporate de novo GWAS are a powerful new tool for tropical rice
improvement. \emph{Heredity}, \emph{116}(4), 395--408.
\url{https://doi.org/10.1038/hdy.2015.113}

\bibitem[\citeproctext]{ref-akdemir2015locally}
Akdemir, D., \& Jannink, J.-L. (2015). Locally epistatic genomic
relationship matrices for genomic association and prediction.
\emph{Genetics}, \emph{199}(3), 857--871.
\url{https://doi.org/10.1534/genetics.114.173658}

\bibitem[\citeproctext]{ref-akdemir2015optimization}
Akdemir, D., Sanchez, J. I., \& Jannink, J.-L. (2015). Optimization of
genomic selection training populations with a genetic algorithm.
\emph{Genetics Selection Evolution}, \emph{47}(1), 1--10.
\url{https://doi.org/10.1186/s12711-015-0116-6}

\bibitem[\citeproctext]{ref-isidro2015training}
Isidro, J., Jannink, J.-L., Akdemir, D., Poland, J., Heslot, N., \&
Sorrells, M. E. (2015). Training set optimization under population
structure in genomic selection. \emph{Theoretical and Applied Genetics},
\emph{128}(1), 145--158. \url{https://doi.org/10.1186/s12711-015-0116-6}

\bibitem[\citeproctext]{ref-spindel2015genomic}
Spindel, J., Begum, H., Akdemir, D., Virk, P., Collard, B., Redona, E.,
\ldots{} McCouch, S. R. (2015). Genomic selection and association
mapping in rice (oryza sativa): Effect of trait genetic architecture,
training population composition, marker number and statistical model on
accuracy of rice genomic selection in elite, tropical rice breeding
lines. \emph{PLoS Genet}, \emph{11}(2), e1004982.
\url{https://doi.org/10.1371/journal.pgen.1005350}

\bibitem[\citeproctext]{ref-heslot2014integrating}
Heslot, N., Akdemir, D., Sorrells, M. E., \& Jannink, J.-L. (2014).
Integrating environmental covariates and crop modeling into the genomic
selection framework to predict genotype by environment interactions.
\emph{Theoretical and Applied Genetics}, \emph{127}(2), 463--480.
\url{https://doi.org/10.2135/cropsci2014.03.0249}

\bibitem[\citeproctext]{ref-yanzeabase}
Yan, A., Edwards, J. D., Menda, N., Tecle, I. Y., Akdemir, D., Tiede,
T., \ldots{} Jannink, J.-L. (2014). Zeabase: A data retrieval web-portal
and bioinformatics platform for maize genomic selection. \emph{BMC
Bioinformatics}, \emph{15}(1).
\url{https://doi.org/10.1186/s12859-014-0398-7}

\bibitem[\citeproctext]{ref-akdemir2011array}
Akdemir, D., \& Gupta, A. K. (2011). Array variate random variables with
multiway kronecker delta covariance matrix structure. \emph{Journal of
Algebraic Statistics}, \emph{2}(1).
\url{https://doi.org/10.18409/jas.v2i1.12}

\bibitem[\citeproctext]{ref-roa2011squamous}
Roa, J. C., Tapia, O., Cakir, A., Basturk, O., Dursun, N., Akdemir, D.,
\ldots{} Adsay, N. V. (2011). Squamous cell and adenosquamous carcinomas
of the gallbladder: Clinicopathological analysis of 34 cases identified
in 606 carcinomas. \emph{Modern Pathology}, \emph{24}(8), 1069--1078.
\url{https://doi.org/10.1038/modpathol.2011.68}

\bibitem[\citeproctext]{ref-akdemir2010matrix}
Akdemir, D., \& Gupta, A. K. (2010). A matrix variate skew distribution.
\emph{European Journal of Pure and Applied Mathematics}, \emph{3}(2),
128--140. \url{https://doi.org/10.18409/jas.v2i1.12}

\bibitem[\citeproctext]{ref-dursun2010intravesicular}
Dursun, N., Roa, J., Tapia, O., Cakir, A., Coban, I., Basturk, O.,
\ldots{} Adsay, N. (2010). Intravesicular papillary-tubular neoplasm
(IVPN) as a unifying category for mass-forming preinvasive neoplasms of
the gallbladder: An analysis of 87 cases: Clinicopathologic and
immunohistochemical analysis of 123 cases. \emph{LABORATORY
INVESTIGATION}, \emph{90}, 144A--144A.
\url{https://doi.org/10.1097/pas.0b013e318262787c}

\bibitem[\citeproctext]{ref-akdemir2003components}
Akdemir, D. (2003). \emph{Components of response variance for cluster
samples} (Master's thesis, Middle East Technical University; Vol. 37,
pp. 9--12). \url{https://doi.org/10.5152/tpd.2013.03}

\end{CSLReferences}

\section{Education}\label{education}

\begin{itemize}
\tightlist
\item
  \textbf{PhD. in Statistics}, Bowling Green State University, Bowling
  Green, OH, USA, 2009.
\item
  \textbf{M.A.~in Applied Statistics}, Bowling Green State University,
  Bowling Green, OH, USA, 2004.
\item
  \textbf{M.S. in Statistics}, Middle East Technical University, Ankara,
  Turkey, 2003.
\item
  \textbf{B.A. in Business Administration}, Middle East Technical
  University, Ankara, Turkey, 1999.
\end{itemize}

\section{Professional References}\label{professional-references}

\begin{itemize}
\tightlist
\item
  \textbf{Dr.~Yung-Tsi Bolon}

  \begin{itemize}
  \tightlist
  \item
    \textbf{Affiliation:} Director, Immunobiology \& Bioinformatics
    Research, NMDP, Minneapolis, Minnesota, United States
  \item
    \textbf{Relationship:} Supervisor at the National Marrow Donor
    Program
  \item
    \textbf{Contact:}
    \href{mailto:ybolon@nmdp.org}{\nolinkurl{ybolon@nmdp.org}}
  \end{itemize}
\item
  \textbf{Dr.~Julio Isidro-Sanchez}

  \begin{itemize}
  \tightlist
  \item
    \textbf{Affiliation:} Associate Professor: Centro de Biotecnologia y
    Genomica de Plantas, Universidad Politecnica de Madrid, Instituto
    Nacional de Investigacion y Tecnologia Agraria y Alimentaria, Campus
    de Montegancedo - UPM, 28223-Pozuelo de Alarcon, Madrid, Spain
  \item
    \textbf{Relationship:} Expert in plant breeding and genetics,
    collaborator on various projects
  \item
    \textbf{Contact:}
    \href{mailto:j.isidro@upm.es}{\nolinkurl{j.isidro@upm.es}}
  \end{itemize}
\item
  \textbf{Dr.~Jhonathan Pedroso}

  \begin{itemize}
  \tightlist
  \item
    \textbf{Affiliation:} Research Scientist at Corteva Agriscience,
    Corteva, Johnston, Iowa, USA
  \item
    \textbf{Relationship:} Industry partner in genomic tool development,
    contributed to software enhancements
  \item
    \textbf{Contact:}
    \href{mailto:jhowpd@gmail.com}{\nolinkurl{jhowpd@gmail.com}}
  \end{itemize}
\item
  \textbf{Dr.~Lynn Johnson}

  \begin{itemize}
  \tightlist
  \item
    \textbf{Affiliation:} Interim Director and Statistical Consultant,
    Cornell Statistical Consulting Unit, Cornell University, Ithaca, NY,
    USA
  \item
    \textbf{Relationship:} Coworker at the Cornell Statistical
    Consulting Unit
  \item
    \textbf{Contact:}
    \href{mailto:lms86@cornell.edu}{\nolinkurl{lms86@cornell.edu}}
  \end{itemize}
\item
  \textbf{Dr.~Roberto Fritsche Neto}

  \begin{itemize}
  \tightlist
  \item
    \textbf{Affiliation:} Assistant Professor, Department of Plant,
    Environmental Management \& Soil Sciences, LSU
  \item
    \textbf{Relationship:} Collaborator on various projects
  \item
    \textbf{Contact:}
    \href{mailto:rneto1@lsu.edu}{\nolinkurl{rneto1@lsu.edu}}
  \end{itemize}
\end{itemize}

\section{Core Skills}\label{core-skills-1}

\begin{itemize}
\tightlist
\item
  \textbf{Statistical Genomics:} Genomic selection, multi-trait
  methodologies, genomic prediction, GWAS, models for epistasis and
  genotype by environment interactions, optimization of plant and animal
  breeding programs.
\item
  \textbf{Statistics:} Multivariate analysis, high-dimensional data
  modeling, Bayesian methods, machine learning, deep learning, data
  science, biostatistics.
\item
  \textbf{Statistical Computation:} Proficient in R, SAS, Python, and
  C++ for statistical programming and analysis.
\item
  \textbf{Software Development:} Development of research and commercial
  grade software.
\item
  \textbf{Management and administration:} Strategic planning, project
  management, team leadership, grant/patent writing, budgeting,
  accounting.
\end{itemize}

\section{Career Summary}\label{career-summary-1}

\begin{itemize}
\tightlist
\item
  \textbf{Senior Clinical Data Scientist, National Marrow Donor Program,
  Minneapolis, USA (2023 - Current):} Engaged in statistical and machine
  learning analysis of stem cell transplant data, focusing on research
  into donor optimization. Applied for grants and submitted manuscripts
  to peer-reviewed journals. Wrote patent applications.
\item
  \textbf{Clinical Data Scientist, National Marrow Donor Program,
  Minneapolis, USA (2021 - 2023):} Engaged in statistical and machine
  learning analysis of stem cell transplant data, focusing on research
  into donor optimization.
\item
  \textbf{Postdoctoral Research Associate, School of Agriculture and
  Food Science, University College Dublin, Dublin, Ireland (2019 -
  2021):} Conducted research on methods for combining heterogeneous
  genomic and phenotypic datasets and prepared statistical software for
  data analysis.
\item
  \textbf{Statistical Consultant, Cornell Statistical Consulting Unit,
  Cornell University, Ithaca, NY, USA (2017 - 2019):} Provided
  statistical consulting services for researchers at Cornell University,
  including the preparation and presentation of statistics workshops.
\item
  \textbf{Postdoctoral Research Associate, Department of Plant Breeding
  and Genetics, Cornell University, Ithaca, NY, USA (2011 - 2017):}
  Focused on research developing new methodologies in genomic selection
  and prediction, mixed models, and machine learning, advising graduate
  students and preparing statistical software.
\item
  \textbf{Visiting Assistant Professor, Department of Statistics and
  Actuarial Science, University of Central Florida, Orlando, FL, USA
  (2010 - 2011):} Responsibilities included teaching Data Mining
  Methodology, Theoretical Statistics, Applied Time Series Analysis, and
  Nonparametric Statistics.
\item
  \textbf{Visiting Assistant Professor, Department of Mathematics and
  Statistics, Ohio Northern University, Ada, OH, USA (2009 - 2010):}
  Taught Statistics for Professionals, Statistics for Engineers, and
  Statistical Computing, catering to various undergraduate levels.
\end{itemize}

\section{Workshops and Training
Sessions}\label{workshops-and-training-sessions-1}

\begin{itemize}
\tightlist
\item
  \textbf{Genomic Assisted Breeding Workshops:} Conducted workshops in
  the USA, Ireland, Spain, and Belgium, educating the agricultural
  community on genomic technologies.
\item
  \textbf{Public Speaking:} Presented at various international
  conferences including EFI conference in Geneva and ASHI Annual Meeting
  in San Antonio.
\end{itemize}

\section{Publications}\label{publications-1}

\phantomsection\label{refs-fc9fcad9798ae22e92c54d59a6d1bc3c}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-akdemir2020combining}
Akdemir, D., Knox, R., \& Isidro y Sanchez, J. (2020). Combining
partially overlapping multi-omics data in databases using relationship
matrices. \emph{Frontiers in Plant Science}, \emph{11}, 947.
\url{https://doi.org/10.3389/fpls.2020.00947}

\bibitem[\citeproctext]{ref-pinho2020using}
Pinho Morais, P. P., Akdemir, D., Braatz de Andrade, L. R., Jannink,
J.-L., Fritsche-Neto, R., Borem, A., \ldots{} Granato, I. S. C. (2020).
Using public databases for genomic prediction of tropical maize lines.
\emph{Plant Breeding}, \emph{139}(4), 697--707.
\url{https://doi.org/10.1111/pbr.12827}

\bibitem[\citeproctext]{ref-akdemir2019multi}
Akdemir, D., Beavis, W., Fritsche-Neto, R., Singh, A. K., \&
Isidro-Sanchez, J. (2019). Multi-objective optimized genomic breeding
strategies for sustainable food improvement. \emph{Heredity},
\emph{122}(5), 672--683.
\url{https://doi.org/10.1038/s41598-018-38081-6}

\bibitem[\citeproctext]{ref-akdemir2019design}
Akdemir, D., \& Isidro-Sanchez, J. (2019). Design of training
populations for selective phenotyping in genomic prediction.
\emph{Scientific Reports}, \emph{9}(1), 1--15.
\url{https://doi.org/10.1038/s41598-018-38081-6}

\bibitem[\citeproctext]{ref-ikeogu2019genetic}
Ikeogu, U. N., Akdemir, D., Wolfe, M. D., Okeke, U. G., Chinedozi, A.,
Jannink, J.-L., \& Egesi, C. N. (2019). Genetic correlation, genome-wide
association and genomic prediction of portable NIRS predicted
carotenoids in cassava roots. \emph{Frontiers in Plant Science},
\emph{10}, 1570. \url{https://doi.org/10.3389/fpls.2019.01570}

\bibitem[\citeproctext]{ref-kaya2019genome}
Kaya, H. B., Akdemir, D., Lozano, R., Cetin, O., Kaya, H. S., Sahin, M.,
\ldots{} Jannink, J.-L. (2019). Genome wide association study of 5
agronomic traits in olive (olea europaea l.). \emph{Scientific Reports},
\emph{9}(1), 1--14. \url{https://doi.org/10.1038/s41598-019-55338-w}

\bibitem[\citeproctext]{ref-fristche2018accuracy}
Fristche-Neto, R., Akdemir, D., \& Jannink, J.-L. (2018). Accuracy of
genomic selection to predict maize single-crosses obtained through
different mating designs. \emph{Theoretical and Applied Genetics},
\emph{131}(5), 1153--1162.
\url{https://doi.org/10.1007/s00122-018-3068-8}

\bibitem[\citeproctext]{ref-horn2018importance}
Horn, L., Shimelis, H., Kariaga, M., Onamu, R., Wakhungu, J., Were, H.,
et al.others. (2018). Importance of cowpea production, breeding and
production constrains under dry areas in africa. \emph{RUFORUM Working
Document Series}, \emph{17}(1), 499--514.
\url{https://doi.org/10.1016/j.aoas.2020.03.002}

\bibitem[\citeproctext]{ref-okeke2018regional}
Okeke, U. G., Akdemir, D., Rabbi, I., Kulakow, P., \& Jannink, J.-L.
(2018). Regional heritability mapping provides insights into dry matter
content in african white and yellow cassava populations. \emph{The Plant
Genome}, \emph{11}(1), 170050.
\url{https://doi.org/10.3835/plantgenome2017.06.0050}

\bibitem[\citeproctext]{ref-akdemir2017locally}
Akdemir, D., Jannink, J.-L., \& Isidro-Sanchez, J. (2017). Locally
epistatic models for genome-wide prediction and association by
importance sampling. \emph{Genetics Selection Evolution}, \emph{49}(1),
1--14. \url{https://doi.org/10.1186/s12711-017-0348-8}

\bibitem[\citeproctext]{ref-isidro2017genome}
Isidro-Sanchez, J., Akdemir, D., \& Montilla-Bascon, G. (2017).
Genome-wide association analysis using r. In \emph{Oat} (pp. 189--207).
\url{https://doi.org/10.1007/978-1-4939-6682-0_14}

\bibitem[\citeproctext]{ref-okeke2017accuracies}
Okeke, U. G., Akdemir, D., Rabbi, I., Kulakow, P., \& Jannink, J.-L.
(2017). Accuracies of univariate and multivariate genomic prediction
models in african cassava. \emph{Genetics Selection Evolution},
\emph{49}(1), 1--10. \url{https://doi.org/10.1101/116301}

\bibitem[\citeproctext]{ref-ratervar2017}
Stansell, Z., Akdemir, D., \& Bjorkman. (2017). Use of a quality trait
index to increase the reliability of phenotypic evaluations.
\emph{HortScience}, \emph{52}, 1490.

\bibitem[\citeproctext]{ref-akdemir2016array}
Akdemir, D. (2016). Array normal model and incomplete array variate
observations. In \emph{Applied matrix and tensor variate data analysis}
(pp. 93--122). \url{https://doi.org/10.1007/978-4-431-55387-8_5}

\bibitem[\citeproctext]{ref-akdemir2016efficient}
Akdemir, D., \& Sanchez, J. I. (2016). Efficient breeding by genomic
mating. \emph{Frontiers in Genetics}, \emph{7}, 210.
\url{https://doi.org/10.3389/fgene.2016.00210}

\bibitem[\citeproctext]{ref-ashraf2016genomic}
Ashraf, B., Edriss, V., Akdemir, D., Autrique, E., Bonnett, D., Crossa,
J., \ldots{} Jannink, J.-L. (2016). Genomic prediction using phenotypes
from pedigreed lines with no marker data. \emph{Crop Science},
\emph{56}(3), 957--964.
\url{https://doi.org/10.2135/cropsci2015.02.0111}

\bibitem[\citeproctext]{ref-crossa2016genomic}
Crossa, J., Jarquin, D., Franco, J., Perez-Rodriguez, P., Burgueno, J.,
Saint-Pierre, C., et al.others. (2016). Genomic prediction of gene bank
wheat landraces. \emph{G3: Genes, Genomes, Genetics}, \emph{6}(7),
1819--1834.

\bibitem[\citeproctext]{ref-jafarzadeh2016breeding}
Jafarzadeh, J., Bonnett, D., Jannink, J.-L., Akdemir, D., Dreisigacker,
S., \& Sorrells, M. E. (2016). Breeding value of primary synthetic wheat
genotypes for grain yield. \emph{PloS One}, \emph{11}(9), e0162860.
\url{https://doi.org/10.1371/journal.pone.0162860}

\bibitem[\citeproctext]{ref-jafarzadehdetermining}
Jafarzadeh, J., Bonnett, D., Jannink, J.-L., Akdemir, D., \& Sorrells,
M. E. (2016). Determining the breeding value of primary synthetic wheats
for increased grain yield. \emph{PLOS ONE}, \emph{11}(9), e0162860.
\url{https://doi.org/10.1371/journal.pone.0162860}

\bibitem[\citeproctext]{ref-spindel2016genome}
Spindel, J., Begum, H., Akdemir, D., Collard, B., Redona, E., Jannink,
J., \& McCouch, S. (2016). Genome-wide prediction models that
incorporate de novo GWAS are a powerful new tool for tropical rice
improvement. \emph{Heredity}, \emph{116}(4), 395--408.
\url{https://doi.org/10.1038/hdy.2015.113}

\bibitem[\citeproctext]{ref-akdemir2015locally}
Akdemir, D., \& Jannink, J.-L. (2015). Locally epistatic genomic
relationship matrices for genomic association and prediction.
\emph{Genetics}, \emph{199}(3), 857--871.
\url{https://doi.org/10.1534/genetics.114.173658}

\bibitem[\citeproctext]{ref-akdemir2015optimization}
Akdemir, D., Sanchez, J. I., \& Jannink, J.-L. (2015). Optimization of
genomic selection training populations with a genetic algorithm.
\emph{Genetics Selection Evolution}, \emph{47}(1), 1--10.
\url{https://doi.org/10.1186/s12711-015-0116-6}

\bibitem[\citeproctext]{ref-isidro2015training}
Isidro, J., Jannink, J.-L., Akdemir, D., Poland, J., Heslot, N., \&
Sorrells, M. E. (2015). Training set optimization under population
structure in genomic selection. \emph{Theoretical and Applied Genetics},
\emph{128}(1), 145--158. \url{https://doi.org/10.1186/s12711-015-0116-6}

\bibitem[\citeproctext]{ref-spindel2015genomic}
Spindel, J., Begum, H., Akdemir, D., Virk, P., Collard, B., Redona, E.,
\ldots{} McCouch, S. R. (2015). Genomic selection and association
mapping in rice (oryza sativa): Effect of trait genetic architecture,
training population composition, marker number and statistical model on
accuracy of rice genomic selection in elite, tropical rice breeding
lines. \emph{PLoS Genet}, \emph{11}(2), e1004982.
\url{https://doi.org/10.1371/journal.pgen.1005350}

\bibitem[\citeproctext]{ref-heslot2014integrating}
Heslot, N., Akdemir, D., Sorrells, M. E., \& Jannink, J.-L. (2014).
Integrating environmental covariates and crop modeling into the genomic
selection framework to predict genotype by environment interactions.
\emph{Theoretical and Applied Genetics}, \emph{127}(2), 463--480.
\url{https://doi.org/10.2135/cropsci2014.03.0249}

\bibitem[\citeproctext]{ref-yanzeabase}
Yan, A., Edwards, J. D., Menda, N., Tecle, I. Y., Akdemir, D., Tiede,
T., \ldots{} Jannink, J.-L. (2014). Zeabase: A data retrieval web-portal
and bioinformatics platform for maize genomic selection. \emph{BMC
Bioinformatics}, \emph{15}(1).
\url{https://doi.org/10.1186/s12859-014-0398-7}

\bibitem[\citeproctext]{ref-akdemir2011array}
Akdemir, D., \& Gupta, A. K. (2011). Array variate random variables with
multiway kronecker delta covariance matrix structure. \emph{Journal of
Algebraic Statistics}, \emph{2}(1).
\url{https://doi.org/10.18409/jas.v2i1.12}

\bibitem[\citeproctext]{ref-roa2011squamous}
Roa, J. C., Tapia, O., Cakir, A., Basturk, O., Dursun, N., Akdemir, D.,
\ldots{} Adsay, N. V. (2011). Squamous cell and adenosquamous carcinomas
of the gallbladder: Clinicopathological analysis of 34 cases identified
in 606 carcinomas. \emph{Modern Pathology}, \emph{24}(8), 1069--1078.
\url{https://doi.org/10.1038/modpathol.2011.68}

\bibitem[\citeproctext]{ref-akdemir2010matrix}
Akdemir, D., \& Gupta, A. K. (2010). A matrix variate skew distribution.
\emph{European Journal of Pure and Applied Mathematics}, \emph{3}(2),
128--140. \url{https://doi.org/10.18409/jas.v2i1.12}

\bibitem[\citeproctext]{ref-dursun2010intravesicular}
Dursun, N., Roa, J., Tapia, O., Cakir, A., Coban, I., Basturk, O.,
\ldots{} Adsay, N. (2010). Intravesicular papillary-tubular neoplasm
(IVPN) as a unifying category for mass-forming preinvasive neoplasms of
the gallbladder: An analysis of 87 cases: Clinicopathologic and
immunohistochemical analysis of 123 cases. \emph{LABORATORY
INVESTIGATION}, \emph{90}, 144A--144A.
\url{https://doi.org/10.1097/pas.0b013e318262787c}

\bibitem[\citeproctext]{ref-akdemir2003components}
Akdemir, D. (2003). \emph{Components of response variance for cluster
samples} (Master's thesis, Middle East Technical University; Vol. 37,
pp. 9--12). \url{https://doi.org/10.5152/tpd.2013.03}

\end{CSLReferences}

\section{Education}\label{education-1}

\begin{itemize}
\tightlist
\item
  \textbf{PhD. in Statistics}, Bowling Green State University, Bowling
  Green, OH, USA, 2009.
\item
  \textbf{M.A.~in Applied Statistics}, Bowling Green State University,
  Bowling Green, OH, USA, 2004.
\item
  \textbf{M.S. in Statistics}, Middle East Technical University, Ankara,
  Turkey, 2003.
\item
  \textbf{B.A. in Business Administration}, Middle East Technical
  University, Ankara, Turkey, 1999.
\end{itemize}

\section{Professional References}\label{professional-references-1}

\begin{itemize}
\tightlist
\item
  \textbf{Dr.~Yung-Tsi Bolon}

  \begin{itemize}
  \tightlist
  \item
    \textbf{Affiliation:} Director, Immunobiology \& Bioinformatics
    Research, NMDP, Minneapolis, Minnesota, United States
  \item
    \textbf{Relationship:} Supervisor at the National Marrow Donor
    Program
  \item
    \textbf{Contact:}
    \href{mailto:ybolon@nmdp.org}{\nolinkurl{ybolon@nmdp.org}}
  \end{itemize}
\item
  \textbf{Dr.~Julio Isidro-Sanchez}

  \begin{itemize}
  \tightlist
  \item
    \textbf{Affiliation:} Associate Professor: Centro de Biotecnologia y
    Genomica de Plantas, Universidad Politecnica de Madrid, Instituto
    Nacional de Investigacion y Tecnologia Agraria y Alimentaria, Campus
    de Montegancedo - UPM, 28223-Pozuelo de Alarcon, Madrid, Spain
  \item
    \textbf{Relationship:} Expert in plant breeding and genetics,
    collaborator on various projects
  \item
    \textbf{Contact:}
    \href{mailto:j.isidro@upm.es}{\nolinkurl{j.isidro@upm.es}}
  \end{itemize}
\item
  \textbf{Dr.~Jhonathan Pedroso}

  \begin{itemize}
  \tightlist
  \item
    \textbf{Affiliation:} Research Scientist at Corteva Agriscience,
    Corteva, Johnston, Iowa, USA
  \item
    \textbf{Relationship:} Industry partner in genomic tool development,
    contributed to software enhancements
  \item
    \textbf{Contact:}
    \href{mailto:jhowpd@gmail.com}{\nolinkurl{jhowpd@gmail.com}}
  \end{itemize}
\item
  \textbf{Dr.~Lynn Johnson}

  \begin{itemize}
  \tightlist
  \item
    \textbf{Affiliation:} Interim Director and Statistical Consultant,
    Cornell Statistical Consulting Unit, Cornell University, Ithaca, NY,
    USA
  \item
    \textbf{Relationship:} Coworker at the Cornell Statistical
    Consulting Unit
  \item
    \textbf{Contact:}
    \href{mailto:lms86@cornell.edu}{\nolinkurl{lms86@cornell.edu}}
  \end{itemize}
\item
  \textbf{Dr.~Roberto Fritsche Neto}

  \begin{itemize}
  \tightlist
  \item
    \textbf{Affiliation:} Assistant Professor, Department of Plant,
    Environmental Management \& Soil Sciences, LSU
  \item
    \textbf{Relationship:} Collaborator on various projects
  \item
    \textbf{Contact:}
    \href{mailto:rneto1@lsu.edu}{\nolinkurl{rneto1@lsu.edu}}
  \end{itemize}
\end{itemize}


\end{document}

%\clearpage\end{CJK*}                              % if you are typesetting your resume in Chinese using CJK; the \clearpage is required for fancyhdr to work correctly with CJK, though it kills the page numbering by making \lastpage undefined
\end{document}


%% end of file `template.tex'.
