---
graphics: yes
name: "Dr. Deniz Akdemir"
position: "Machine Learning & AI Specialist with Genomic Expertise"
address: ""
phone: "+1 607 262 6875 (USA) - +34 674 771 859 (Spain)"
email: "deniz.akdemir.work@gmail.com"
github: "denizakdemir"
date: "`r format(Sys.Date(), '%B %Y')`"
output: 
  vitae::moderncv:
    keep_tex: true
csl: "../csl/apa6-cv.csl"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```

# Professional Summary

I am a dedicated machine learning and AI professional with a strong background in statistical genomics and data science. My expertise encompasses designing cutting-edge algorithms, building scalable software solutions, and leading teams in developing advanced applications across genomics and other high-impact fields.

# Key Areas of Expertise

- **Statistical Genomics & Advanced Modeling:**
  - Genomic selection, multi-trait approaches, GWAS, and epistasis modeling
  - Integration of genotype-by-environment interactions into AI models
  - Optimization frameworks for plant and animal breeding

- **Machine Learning & Deep Learning:**
  - End-to-end modeling pipelines for supervised and unsupervised learning
  - High-dimensional data handling, dimensionality reduction, feature engineering
  - Bayesian deep learning, ensemble methods, and model interpretability

- **Statistical & Computational Skills:**
  - Programming in R, SAS, Python, and C++ for algorithm design and deployment
  - Development of computationally efficient, scalable AI solutions
  - Deployment of machine learning models in research and production environments

- **Software Development & Architecture:**
  - Building robust, modular software with version control and CI/CD integration
  - API design, cloud computing, and containerization for scalable services

- **Project Management & Leadership:**
  - Leading multidisciplinary AI projects and managing cross-functional teams
  - Grant writing, patent drafting, and managing budgets for innovative research


# Career Summary
-   **Senior Clinical Data Scientist, National Marrow Donor Program, Minneapolis, USA (2023 - Current):** Engaged in statistical and machine learning analysis of stem cell transplant data, focusing on research into donor optimization. Applied for grants and submitted manuscripts to peer-reviewed journals. Wrote patent applications.
-   **Clinical Data Scientist, National Marrow Donor Program, Minneapolis, USA (2021 - 2023):** Engaged in statistical and machine learning analysis of stem cell transplant data, focusing on research into donor optimization.
- **Postdoctoral Research Associate, School of Agriculture and Food Science, University College Dublin, Dublin, Ireland (2019 - 2021):** Conducted research on methods for combining heterogeneous genomic and phenotypic datasets and prepared statistical software for data analysis.
- **Statistical Consultant, Cornell Statistical Consulting Unit, Cornell University, Ithaca, NY, USA (2017 - 2019):** Provided statistical consulting services for researchers at Cornell University, including the preparation and presentation of statistics workshops.
- **Postdoctoral Research Associate, Department of Plant Breeding and Genetics, Cornell University, Ithaca, NY, USA (2011 - 2017):** Focused on research developing new methodologies in genomic selection and prediction, mixed models, and machine learning, advising graduate students and preparing statistical software.
- **Visiting Assistant Professor, Department of Statistics and Actuarial Science, University of Central Florida, Orlando, FL, USA (2010 - 2011):** Responsibilities included teaching Data Mining Methodology, Theoretical Statistics, Applied Time Series Analysis, and Nonparametric Statistics.
- **Visiting Assistant Professor, Department of Mathematics and Statistics, Ohio Northern University, Ada, OH, USA (2009 - 2010):** Taught Statistics for Professionals, Statistics for Engineers, and Statistical Computing, catering to various undergraduate levels.

# Workshops and Training Sessions

- **Genomic Assisted Breeding Workshops:** Conducted workshops in the USA, Ireland, Spain, and Belgium, educating the agricultural community on genomic technologies.
- **Public Speaking:** Presented at various international conferences including EFI conference in Geneva and ASHI Annual Meeting in San Antonio.

# Publications

```{r references, echo=FALSE, as.is=TRUE, warning=FALSE, message=FALSE}
# Read all publications
all_publications <- vitae::bibliography_entries("../../data/DA_bibliography.bib")

# Filter for AI/ML-related publications
ai_ml_keywords <- c("machine learning", "deep learning", "artificial intelligence", "neural network", 
                    "prediction", "algorithm", "optimization", "data science", "bayesian", 
                    "ensemble", "modeling", "predictive", "computational")
combined_text <- paste(
  tolower(all_publications$title), 
  tolower(as.character(all_publications$journal)),
  sep = " "
)
ai_ml_publications <- all_publications[grepl(paste(ai_ml_keywords, collapse="|"), 
                                           combined_text, 
                                           fixed = FALSE)]

# Display AI/ML-focused publications (only if found, otherwise show first 10)
if(length(ai_ml_publications) > 0) {
  ai_ml_publications
} else {
  # Fallback to show first 10 publications
  all_publications[1:min(10, length(all_publications))]
}
```

# Education

- **PhD. in Statistics**, Bowling Green State University, Bowling Green, OH, USA, 2009.
- **M.A. in Applied Statistics**, Bowling Green State University, Bowling Green, OH, USA, 2004.
- **M.S. in Statistics**, Middle East Technical University, Ankara, Turkey, 2003.
- **B.A. in Business Administration**, Middle East Technical University, Ankara, Turkey, 1999.

# Professional References

- **Dr. Yung-Tsi Bolon**
  - **Affiliation:** Director, Immunobiology & Bioinformatics Research, NMDP, Minneapolis, Minnesota, United States
  - **Relationship:** Supervisor at the National Marrow Donor Program
  - **Contact:** ybolon@nmdp.org
- **Dr. Julio Isidro-Sanchez**
  - **Affiliation:** Associate Professor: Centro de Biotecnologia y Genomica de Plantas, Universidad Politecnica de Madrid, Instituto Nacional de Investigacion y Tecnologia Agraria y Alimentaria, Campus de Montegancedo - UPM, 28223-Pozuelo de Alarcon, Madrid, Spain
  - **Relationship:** Expert in plant breeding and genetics, collaborator on various projects
  - **Contact:** j.isidro@upm.es
- **Dr. Jhonathan Pedroso**
  - **Affiliation:** Research Scientist at Corteva Agriscience, Corteva, Johnston, Iowa, USA
  - **Relationship:** Industry partner in genomic tool development, contributed to software enhancements
  - **Contact:** jhowpd@gmail.com
- **Dr. Lynn Johnson**
  - **Affiliation:** Interim Director and Statistical Consultant, Cornell Statistical Consulting Unit, Cornell University, Ithaca, NY, USA
  - **Relationship:** Coworker at the Cornell Statistical Consulting Unit
  - **Contact:** lms86@cornell.edu
- **Dr. Roberto Fritsche Neto**
  - **Affiliation:** Assistant Professor, Department of Plant, Environmental Management & Soil Sciences, LSU
  - **Relationship:** Collaborator on various projects
  - **Contact:** rneto1@lsu.edu
