---
graphics: yes
name: "Dr. Deniz Akdemir"
position: "Data Science & Statistical Genomics Professional"
address: ""
phone: "+1 607 262 6875"
email: "deniz.akdemir.work@gmail.com"
github: "denizakdemir"
date: "`r format(Sys.Date(), '%B %Y')`"
output: 
  vitae::moderncv:
    keep_tex: true
    includes:
      in_header: ../../styles/header_include.tex
csl: "../../csl/apa6-cv.csl"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```

# Professional Summary

I am a seasoned data scientist and statistical genomics professional with a versatile skill set spanning statistics, machine learning, and software development. Through effective management and cross-functional teamwork, I bring academic rigor and commercial insight to complex data challenges, ensuring both high-quality results and impactful project outcomes.

# Key Areas of Expertise

- **Statistical Genomics:**
  - Comprehensive knowledge of genomic selection, GWAS, and epistasis
  - Leveraging multi-trait models and genotype-by-environment interactions
  - Enhancing breeding programs through advanced data-driven strategies

- **Statistics & Data Analysis:**
  - Multivariate analysis, high-dimensional modeling, and Bayesian inference
  - Machine learning and deep learning for predictive analytics
  - Biostatistics applications in research and industry

- **Technical Proficiencies:**
  - Skilled in R, SAS, Python, and C++ for robust statistical computing
  - Development of both research and commercial-grade software

- **Leadership & Management:**
  - Strong project management, strategic planning, and budget oversight
  - Team leadership with a focus on innovation and collaborative research
  - Proven track record in grant writing and intellectual property management


# Career Summary
-   **Senior Clinical Data Scientist, National Marrow Donor Program, Minneapolis, USA (2023 - Current):** Engaged in statistical and machine learning analysis of stem cell transplant data, focusing on research into donor optimization. Applied for grants and submitted manuscripts to peer-reviewed journals. Wrote patent applications.
-   **Clinical Data Scientist, National Marrow Donor Program, Minneapolis, USA (2021 - 2023):** Engaged in statistical and machine learning analysis of stem cell transplant data, focusing on research into donor optimization.
- **Postdoctoral Research Associate, School of Agriculture and Food Science, University College Dublin, Dublin, Ireland (2019 - 2021):** Conducted research on methods for combining heterogeneous genomic and phenotypic datasets and prepared statistical software for data analysis.
- **Statistical Consultant, Cornell Statistical Consulting Unit, Cornell University, Ithaca, NY, USA (2017 - 2019):** Provided statistical consulting services for researchers at Cornell University, including the preparation and presentation of statistics workshops.
- **Postdoctoral Research Associate, Department of Plant Breeding and Genetics, Cornell University, Ithaca, NY, USA (2011 - 2017):** Focused on research developing new methodologies in genomic selection and prediction, mixed models, and machine learning, advising graduate students and preparing statistical software.
- **Visiting Assistant Professor, Department of Statistics and Actuarial Science, University of Central Florida, Orlando, FL, USA (2010 - 2011):** Responsibilities included teaching Data Mining Methodology, Theoretical Statistics, Applied Time Series Analysis, and Nonparametric Statistics.
- **Visiting Assistant Professor, Department of Mathematics and Statistics, Ohio Northern University, Ada, OH, USA (2009 - 2010):** Taught Statistics for Professionals, Statistics for Engineers, and Statistical Computing, catering to various undergraduate levels.

# Workshops and Training Sessions

- **Genomic Assisted Breeding Workshops:** Conducted workshops in the USA, Ireland, Spain, and Belgium, educating the agricultural community on genomic technologies.
- **Public Speaking:** Presented at various international conferences including EFI conference in Geneva and ASHI Annual Meeting in San Antonio.

# Publications

```{r references, echo=FALSE, as.is=TRUE, warning=FALSE, message=FALSE}
vitae::bibliography_entries("../../data/DA_bibliography.bib")
```

# Education

- **PhD. in Statistics**, Bowling Green State University, Bowling Green, OH, USA, 2009.
- **M.A. in Applied Statistics**, Bowling Green State University, Bowling Green, OH, USA, 2004.
- **M.S. in Statistics**, Middle East Technical University, Ankara, Turkey, 2003.
- **B.A. in Business Administration**, Middle East Technical University, Ankara, Turkey, 1999.

# Professional References

- **Dr. Yung-Tsi Bolon**
  - **Affiliation:** Director, Immunobiology & Bioinformatics Research, NMDP, Minneapolis, Minnesota, United States
  - **Relationship:** Supervisor at the National Marrow Donor Program
  - **Contact:** ybolon@nmdp.org
- **Dr. Julio Isidro-Sanchez**
  - **Affiliation:** Associate Professor: Centro de Biotecnologia y Genomica de Plantas, Universidad Politecnica de Madrid, Instituto Nacional de Investigacion y Tecnologia Agraria y Alimentaria, Campus de Montegancedo - UPM, 28223-Pozuelo de Alarcon, Madrid, Spain
  - **Relationship:** Expert in plant breeding and genetics, collaborator on various projects
  - **Contact:** j.isidro@upm.es
- **Dr. Jhonathan Pedroso**
  - **Affiliation:** Research Scientist at Corteva Agriscience, Corteva, Johnston, Iowa, USA
  - **Relationship:** Industry partner in genomic tool development, contributed to software enhancements
  - **Contact:** jhowpd@gmail.com
- **Dr. Lynn Johnson**
  - **Affiliation:** Interim Director and Statistical Consultant, Cornell Statistical Consulting Unit, Cornell University, Ithaca, NY, USA
  - **Relationship:** Coworker at the Cornell Statistical Consulting Unit
  - **Contact:** lms86@cornell.edu
- **Dr. Roberto Fritsche Neto**
  - **Affiliation:** Assistant Professor, Department of Plant, Environmental Management & Soil Sciences, LSU
  - **Relationship:** Collaborator on various projects
  - **Contact:** rneto1@lsu.edu
