---
graphics: yes
name: "Dr. Deniz Akdemir"
position: "Strategic Leadership in Data Science & Genomics"
address: ""
phone: "+1 607 262 6875"
email: "deniz.akdemir.work@gmail.com"
github: "denizakdemir"
date: "`r format(Sys.Date(), '%B %Y')`"
output: 
  vitae::moderncv:
    keep_tex: true
    includes:
      in_header: ../../styles/header_include.tex
csl: "../../csl/apa6-cv.csl"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```

# Executive Profile

I am a strategic leader with deep technical expertise in data science, statistical genomics, and computational biology. With a proven track record of driving innovation through effective team leadership and cross-functional collaboration, I excel at transforming complex scientific challenges into practical, impactful solutions. My approach combines rigorous methodological standards with business acumen to deliver measurable results in both academic and commercial settings.

# Leadership Competencies

- **Strategic Vision & Execution:**
  - Translating organizational objectives into actionable research plans
  - Aligning technical innovation with business priorities
  - Delivering complex scientific projects on time and within budget

- **Team Leadership & Development:**
  - Building and mentoring high-performing multidisciplinary teams
  - Fostering collaborative environments that maximize team productivity
  - Developing talent and creating career growth opportunities for team members

- **Resource Management:**
  - Securing and managing grant funding and commercial research budgets
  - Optimizing allocation of personnel and technical resources
  - Strategic planning for sustainable research and technology development

- **Stakeholder Management:**
  - Effectively communicating technical concepts to diverse audiences
  - Building productive partnerships across academic and industry boundaries
  - Bridging the gap between scientific innovation and practical application

- **Innovation Management:**
  - Patent development and intellectual property strategy
  - Identifying commercial opportunities from scientific breakthroughs
  - Creating frameworks for sustainable scientific innovation

# Career Highlights

- **Led multidisciplinary research teams** across international research institutions, managing collaborative projects with budgets exceeding $1M
- **Developed patentable methodologies** in genomic optimization with commercial applications
- **Secured multiple competitive grants** through effective proposal development and presentation
- **Published over 30 peer-reviewed papers** in high-impact journals, demonstrating scientific leadership
- **Established international partnerships** between academic institutions and commercial enterprises

# Leadership Experience

-   **Senior Clinical Data Scientist, National Marrow Donor Program, Minneapolis, USA (2023 - Current):** 
    - Lead strategic initiatives in donor optimization using advanced analytics
    - Develop and implement novel machine learning approaches for clinical data
    - Oversee grant development and patent applications for innovative methodologies
    - Collaborate with cross-functional teams to translate research into clinical applications

-   **Clinical Data Scientist, National Marrow Donor Program, Minneapolis, USA (2021 - 2023):** 
    - Directed statistical and machine learning analyses of stem cell transplant data
    - Coordinated with clinical teams to ensure research relevance to medical applications
    - Managed data science workflows and quality control processes
    - Presented research findings to executive stakeholders

- **Postdoctoral Research Associate, School of Agriculture and Food Science, University College Dublin, Dublin, Ireland (2019 - 2021):** 
    - Led research on methods for combining heterogeneous genomic and phenotypic datasets
    - Supervised junior researchers and graduate students
    - Developed international research collaborations
    - Secured additional project funding through grant applications

- **Statistical Consultant, Cornell Statistical Consulting Unit, Cornell University, Ithaca, NY, USA (2017 - 2019):** 
    - Managed statistical consulting services for diverse research departments
    - Developed and presented statistics workshops to faculty and students
    - Recommended methodological approaches for complex research questions
    - Ensured statistical rigor in university-wide research initiatives

# Strategic Initiatives

- **Genomic Optimization Platform:** Led the conceptual development and implementation of a computational framework for optimizing breeding programs, resulting in improved efficiency and outcomes.

- **Multi-institutional Collaboration Network:** Established and coordinated research partnerships across 5 countries, enabling resource sharing and accelerated research outcomes.

- **Integrated Data Science Pipeline:** Developed a comprehensive workflow for integrating heterogeneous data sources, enhancing predictive capabilities and research insights.

- **Workshop Series Development:** Created and implemented specialized training programs for researchers and practitioners, building capacity in advanced genomic technologies.

# Education

- **PhD. in Statistics**, Bowling Green State University, Bowling Green, OH, USA, 2009.
- **M.A. in Applied Statistics**, Bowling Green State University, Bowling Green, OH, USA, 2004.
- **M.S. in Statistics**, Middle East Technical University, Ankara, Turkey, 2003.
- **B.A. in Business Administration**, Middle East Technical University, Ankara, Turkey, 1999.

# Select Publications

```{r references, echo=FALSE, as.is=TRUE, warning=FALSE, message=FALSE}
# Read all publications
all_publications <- vitae::bibliography_entries("../../data/DA_bibliography.bib")

# Get first-author publications (likely where you led the work)
first_author_pattern <- "^Akdemir, D"
first_author_pubs <- all_publications[grepl(first_author_pattern, all_publications$author)]

# If we found first-author pubs, use them, otherwise use most recent 5
if(length(first_author_pubs) > 0) {
  # Sort by date if possible and take most recent 5
  if("date" %in% names(first_author_pubs)) {
    sorted_pubs <- first_author_pubs[order(first_author_pubs$date, decreasing=TRUE)]
    leader_pubs <- sorted_pubs[1:min(5, length(sorted_pubs))]
  } else {
    # If can't sort by date, just take first 5
    leader_pubs <- first_author_pubs[1:min(5, length(first_author_pubs))]
  }
} else {
  # Fallback to first 5 publications
  leader_pubs <- all_publications[1:min(5, length(all_publications))]
}

# Display leadership-related publications
leader_pubs
```

# Professional Network

- **Dr. Yung-Tsi Bolon**
  - **Position:** Director, Immunobiology & Bioinformatics Research, NMDP
  - **Collaboration Area:** Strategic development of genomic applications in clinical settings

- **Dr. Julio Isidro-Sanchez**
  - **Position:** Associate Professor, Universidad Politecnica de Madrid
  - **Collaboration Area:** International research partnerships and funding initiatives

- **Dr. Jhonathan Pedroso**
  - **Position:** Research Scientist at Corteva Agriscience
  - **Collaboration Area:** Industry-academic partnerships and technology transfer

- **Dr. Lynn Johnson**
  - **Position:** Interim Director, Cornell Statistical Consulting Unit
  - **Collaboration Area:** Statistical methodology and education initiatives

- **Dr. Roberto Fritsche Neto**
  - **Position:** Assistant Professor, Department of Plant, Environmental Management & Soil Sciences, LSU
  - **Collaboration Area:** Multi-institutional research programs and grant development