---
graphics: yes
name: "Dr. Deniz Akdemir"
position: "Statistical Genomics Specialist"
address: ""
phone: "+1 607 262 6875"
email: "deniz.akdemir.work@gmail.com"
github: "denizakdemir"
date: "`r format(Sys.Date(), '%B %Y')`"
output: 
  vitae::moderncv:
    keep_tex: true
    includes:
      in_header: ../../styles/header_include.tex
csl: "../../csl/apa6-cv.csl"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```

# Professional Summary

I am an accomplished statistical genomics professional with deep expertise in developing advanced methodologies for genomic prediction and selection. My work focuses on optimizing breeding programs and creating computational tools that integrate complex genomic, phenotypic, and environmental data to enhance agricultural and biomedical outcomes.

# Key Areas of Expertise

- **Statistical Genomics:**
  - Genomic selection, multi-trait methodologies, genomic prediction
  - Genome-Wide Association Studies (GWAS)
  - Epistatic models and genotype-by-environment interaction analysis
  - Optimization of plant and animal breeding programs

- **Statistics & Biostatistics:**
  - Multivariate analysis and high-dimensional data modeling
  - Bayesian methods, frequentist approaches, and robust inference
  - Machine learning and deep learning in biological research
  - Applied data science and experimental design

- **Computational Genomics:**
  - Development of software tools for analyzing genomic data
  - Tools for genotype imputation, population structure analysis, and genomic relationships
  - Multi-omics integration of heterogeneous data for improved prediction

- **Applied Genomics:**
  - Practical applications in plant breeding (rice, wheat, cassava, olive)
  - Clinical applications in stem cell transplantation
  - Implementation of genomic selection in commercial breeding programs

- **Leadership & Management:**
  - Strategic planning and project management for genomic initiatives
  - Team leadership and cross-disciplinary collaboration
  - Grant writing and intellectual property development


# Career Summary
-   **Senior Clinical Data Scientist, National Marrow Donor Program, Minneapolis, USA (2023 - Current):** Engaged in statistical and machine learning analysis of stem cell transplant data, focusing on research into donor optimization. Applied for grants and submitted manuscripts to peer-reviewed journals. Wrote patent applications.
-   **Clinical Data Scientist, National Marrow Donor Program, Minneapolis, USA (2021 - 2023):** Engaged in statistical and machine learning analysis of stem cell transplant data, focusing on research into donor optimization.
- **Postdoctoral Research Associate, School of Agriculture and Food Science, University College Dublin, Dublin, Ireland (2019 - 2021):** Conducted research on methods for combining heterogeneous genomic and phenotypic datasets and prepared statistical software for data analysis.
- **Statistical Consultant, Cornell Statistical Consulting Unit, Cornell University, Ithaca, NY, USA (2017 - 2019):** Provided statistical consulting services for researchers at Cornell University, including the preparation and presentation of statistics workshops.
- **Postdoctoral Research Associate, Department of Plant Breeding and Genetics, Cornell University, Ithaca, NY, USA (2011 - 2017):** Focused on research developing new methodologies in genomic selection and prediction, mixed models, and machine learning, advising graduate students and preparing statistical software.
- **Visiting Assistant Professor, Department of Statistics and Actuarial Science, University of Central Florida, Orlando, FL, USA (2010 - 2011):** Responsibilities included teaching Data Mining Methodology, Theoretical Statistics, Applied Time Series Analysis, and Nonparametric Statistics.
- **Visiting Assistant Professor, Department of Mathematics and Statistics, Ohio Northern University, Ada, OH, USA (2009 - 2010):** Taught Statistics for Professionals, Statistics for Engineers, and Statistical Computing, catering to various undergraduate levels.

# Workshops and Training Sessions

- **Genomic Assisted Breeding Workshops:** Conducted workshops in the USA, Ireland, Spain, and Belgium, educating the agricultural community on genomic technologies.
- **Public Speaking:** Presented at various international conferences including EFI conference in Geneva and ASHI Annual Meeting in San Antonio.

# Publications

```{r references, echo=FALSE, as.is=TRUE, warning=FALSE, message=FALSE}
# Read all publications
all_publications <- vitae::bibliography_entries("../../data/DA_bibliography.bib")

# Filter for genomics-related publications
genomics_keywords <- c("genom", "breed", "genetic", "selection", "prediction", "GWAS", "epistasis", 
                       "QTL", "sequenc", "gene", "haplotype", "SNP", "marker", "crop", "plant",
                       "rice", "wheat", "cassava", "olive")
combined_text <- paste(
  tolower(all_publications$title), 
  tolower(as.character(all_publications$journal)),
  sep = " "
)
genomics_publications <- all_publications[grepl(paste(genomics_keywords, collapse="|"), 
                                              combined_text, 
                                              fixed = FALSE)]

# Display genomics-focused publications (only if found, otherwise show first 10)
if(length(genomics_publications) > 0) {
  genomics_publications
} else {
  # Fallback to show first 10 publications
  all_publications[1:min(10, length(all_publications))]
}
```

# Education

- **PhD. in Statistics**, Bowling Green State University, Bowling Green, OH, USA, 2009.
- **M.A. in Applied Statistics**, Bowling Green State University, Bowling Green, OH, USA, 2004.
- **M.S. in Statistics**, Middle East Technical University, Ankara, Turkey, 2003.
- **B.A. in Business Administration**, Middle East Technical University, Ankara, Turkey, 1999.

# Professional References

- **Dr. Yung-Tsi Bolon**
  - **Affiliation:** Director, Immunobiology & Bioinformatics Research, NMDP, Minneapolis, Minnesota, United States
  - **Relationship:** Supervisor at the National Marrow Donor Program
  - **Contact:** ybolon@nmdp.org
- **Dr. Julio Isidro-Sanchez**
  - **Affiliation:** Associate Professor: Centro de Biotecnologia y Genomica de Plantas, Universidad Politecnica de Madrid, Instituto Nacional de Investigacion y Tecnologia Agraria y Alimentaria, Campus de Montegancedo - UPM, 28223-Pozuelo de Alarcon, Madrid, Spain
  - **Relationship:** Expert in plant breeding and genetics, collaborator on various projects
  - **Contact:** j.isidro@upm.es
- **Dr. Jhonathan Pedroso**
  - **Affiliation:** Research Scientist at Corteva Agriscience, Corteva, Johnston, Iowa, USA
  - **Relationship:** Industry partner in genomic tool development, contributed to software enhancements
  - **Contact:** jhowpd@gmail.com
- **Dr. Lynn Johnson**
  - **Affiliation:** Interim Director and Statistical Consultant, Cornell Statistical Consulting Unit, Cornell University, Ithaca, NY, USA
  - **Relationship:** Coworker at the Cornell Statistical Consulting Unit
  - **Contact:** lms86@cornell.edu
- **Dr. Roberto Fritsche Neto**
  - **Affiliation:** Assistant Professor, Department of Plant, Environmental Management & Soil Sciences, LSU
  - **Relationship:** Collaborator on various projects
  - **Contact:** rneto1@lsu.edu
